---
title: "Kernel Density Plot"
author: "lindbrook"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Kernel Density Plot}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(cholera)
```

## Overview

By default, the addKernelDensity() function pools all observations:

```{r fig.width = 5, fig.height = 5, fig.align = "center", echo = TRUE}
snowMap(add.title = FALSE)
addKernelDensity()
```

However, with this function you can also specify which pumps should be considered (i.e., define the "population" of pump neighborhoods) by using the "pump.select" argument and which subset of selected neighborhoods should be plotted by using the "pump.subset" argument.

## Selecting a subset of observed neighborhoods:

```{r fig.width = 5, fig.height = 5, fig.align = "center", echo = TRUE}
snowMap(add.title = FALSE)
addKernelDensity(pump.subset = c(6, 8))
```

## Defining the set of observed pump neighborhoods:

```{r fig.width = 5, fig.height = 5, fig.align = "center", echo = TRUE}
snowMap(add.title = FALSE)
addKernelDensity(pump.select = c(6, 8))
```
