---
title: "Kernel Density Plot"
author: "lindbrook"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Kernel Density Plot}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = ">")
library(cholera)
```
By default, `addKernelDensity()` pools all observations into a single neighborhood:

```{r fig.width = 5, fig.height = 5, fig.align = "center"}
snowMap()
addKernelDensity()
```

To consider the possibility of multiple pump neighborhoods, you can do the following. 

To define a set of neighborhoods based on a selection of pumps (i.e., distribute cases across the selection), use the `pump.select` argument :

```{r fig.width = 5, fig.height = 5, fig.align = "center"}
snowMap()
addKernelDensity(pump.select = c(6, 8))
```

To view the subset of "local" neighborhoods you're interested in, use the `pump.subset` argument.

```{r fig.width = 5, fig.height = 5, fig.align = "center"}
snowMap()
addKernelDensity(pump.subset = c(6, 8))
```
