---
title: "Kernel Density Plot"
author: "lindbrook"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Kernel Density Plot}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(cholera)
```

## Overview

By default, the `addKernelDensity()` function pools all observations:

```{r fig.width = 5, fig.height = 5, fig.align = "center", echo = TRUE}
snowMap()
addKernelDensity()
```

However, the function allows you to explore different scenarios.

## Defining the set of observed pump neighborhoods

By using the `pump.select` argument you can specify which pumps should be considered (i.e., define the "population" of pump neighborhoods).

```{r fig.width = 5, fig.height = 5, fig.align = "center", echo = TRUE}
snowMap()
addKernelDensity(pump.select = c(6, 8))
```

## Selecting a subset of observed neighborhoods

By using the `pump.subset` argument, you can specify which subset of selected neighborhoods will be plotted.

```{r fig.width = 5, fig.height = 5, fig.align = "center", echo = TRUE}
snowMap()
addKernelDensity(pump.subset = c(6, 8))
```
