---
title: "Parallelization"
author: "lindbrook"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Parallelization}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = ">"
)
```

The 'cholera' package supports parallelization via the 'parallel' R package. It is _off_ by default. To enable it, either set `multi.core = TRUE`, which will use all logical cores, or pass the number of cores to use to the argument (e.g., `multi.core = 2`). To check the number of physical/logical cores on your system, use `parallel::detectCores()`. 

Two things to note. First, support for Windows is currently limited. Second, the documentation for the 'parallel' package discourages the use of parallelization in the GUI: "It is strongly discouraged to use these functions in GUI or embedded environments, because it leads to several processes sharing the same GUI which will likely cause chaos (and possibly crashes)." That said, with recent versions of 'parallel', I have rarely, if ever experience crashes in either the R application or RStudio.

Besides courtesy, parallelization is off because revised coding has significantly improved performance for computationally intensive tasks, e.g., area polygons for "expected" cases with neighborhoodWalking() and plot.walking().

## Benchmarks

The tables below compare the timings for 'cholera' versions 0.8.0 and 0.9.0 for the following expressions:

```{r code, eval = FALSE}
plot(neighborhoodWalking())
plot(neighborhoodWalking(case.set = "expected"), type = "area.polygons")
plot(neighborhoodWalking(pump.select = 6:7, case.set = "expected"), type = "area.polygons")
```

The timings (in seconds) reflect the average of 5 evaluations of the code above using the 'microbenchmark' package, R version 4.4.3 and macOS 13.7.4 on a 3.1 GHz Dual-Core Intel Core i5 processor.

| ver. 0.8.0 | 1 core | 4 cores |
:------------- | -------------: | -------------:
observed | 4.0 | 3.9
expected; area.polygons | 285.5 | 131.5
expected; pumps 6 & 7; area.polygons | 227.5 | 121.2 

| ver. 0.9.0 | 1 core | 4 cores |
:------------- | -------------: | -------------:
observed | 2.0 | 2.1
expected; area.polygons | 29.1 | 20.8
expected; pumps 6 & 7; area.polygons | 80.3 | 57.8 
