---
title: "Time Series"
author: "lindbrook"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Time Series}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(cholera)
```

## Overview

The standard "textbook" story is that John Snow helped stem the tide of the outbreak by getting the local government to remove the Broad Street pump's handle.^[https://www.cdc.gov/mmwr/preview/mmwrhtml/mm5334a1.htm].

While Snow likely believed that this would have a positive effect, he himself was skeptical of its actual effect: he writes that by the time the pump was removed on Friday, 08 September 1854, "the epidemic had evidently subsided" (Snow 1855b, 153).

As part of the Vestry Report, both he and the committee included separate tables that appear to reflect separate data that record the number of fatalities. I have included both as separate data sets, "time.series.snow" and "time.series.vestry", in this package.

The data are plotted below. With both sets of data, we can see that the removal of the handle comes after the a significant decline in the fatality level.

## Figures

```{r, fig.width = 5.25, fig.height = 5.25, echo = FALSE}
plot(time.series.vestry$date, time.series.vestry$fatalities, type = "o",
  xlab = "Date", ylab = "fatalities")
abline(v = as.Date("1854-09-08"), col = "red", lty = "dotted")
text(time.series.vestry[time.series.vestry$date == "1854-09-08",
  c("date", "fatalities")], label = "Sep 08", cex = 0.75, col = "red", pos = 4)
title(main = "Vestry Report: Removal of the Broad Street\nPump Handle")

```

```{r, fig.width = 5.25, fig.height = 5.25, echo = FALSE}
plot(time.series.snow$date, time.series.snow$deaths, type = "o", xlab = "Date",
  ylab = "fatalities", xlim = range(time.series.vestry$date), ylim = range(time.series.vestry$fatalities))
abline(v = as.Date("1854-09-08"), col = "red", lty = "dotted")
text(time.series.snow[time.series.snow$date == "1854-09-08",
  c("date", "deaths")], label = "Sep 08", cex = 0.75, col = "red", pos = 4)
title(main = "Snow: Removal of the Broad Street Pump Handle")
```
