---
title: "computing Voronoi diagrams with geographic data"
output:
  # pdf_document: default
  # html_document: default
  output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(cholera)
library(geosphere)
library(terra)
```

## introduction

This is a work-in-progress note about how I compute Voronoi diagrams with geographic data (data with longitude and latitude). The problem is that you can't directly apply the standard Voronoi algorithm, e.g., deldir::deldir(), to such data. Doing so will give you the "wrong" answer.

Part of the reason why is that even though the Earth is a 3D object and its surface is curved, the standard algorithm expects the "world" to be 2D flat. While there are [spatial](https://rspatial.org/) and GIS-style tools and methods that one can use to address this, I'm currently hesitant to adopt them. While doing so may be the wiser long term strategy, right now I'm simply going to try to extend the functionality of ['cholera'](https://github.com/lindbrook/cholera) to include georeferenced versions of the package's data. My strategy for addressing this problem is a familiar one: transform the data so that you _can_ use the algorithm; translate the results back to the original form (longitude and latitude).

If any of the below seems wrong (be it obvious or subtle, technical or conceptual), feel free to share your comments and insights.

## what are Voronoi diagrams?

The best way to understand what a Voronoi diagram is and how it works is through an example. Consider the coffee shops in your town. Let's assume that "all else being equal" (price, quality, banter, etc.) and that the only thing that affects your choice is proximity. If you were an algorithm, you'd compute the distance to each coffee shop and adopt the closest one as your preferred choice. Now, if everyone in town were to do the same, distinct "neighborhoods", based on people's proximity to a shop, will emerge. In this way, the coffee shops would carve up your town.

Here's a hypothetical example of a town where 13 coffee shops (blue triangles) create 13 "neighborhoods" (polygons or cells):

```{r voronoi, echo = FALSE, fig.align = "center"}
snowMap(add.cases = FALSE)
addVoronoi()
title(main = "Figure 1 Voronoi Diagram")
```

If you lived within the boundaries of a given cell then that cell's coffee shop will be the one that's closest to you If you lived on a boundary between two neighborhoods (a cell edge) then you'd happily go to either of those neighborhoods' shop because they'd be equally distant to your home.

## the data

The data in Figure 1 actually come from a digitization of John Snow's map of the 1854 London cholera outbreak done by Dodson and Tobler (1992). So instead of coffee houses, we're looking at water pumps, and instead of a hypothetical town, we're looking at the streets of Soho.

Dodson and Tobler applied a set of Cartesian coordinates to the data. I'll refer to data that use these coordinate as the _nominal_ data. I estimated longitude and latitude (i.e., georeferenced) of these data by using [QGIS](https://qgis.org/) and its `Georeferencer` tool and its [OpenStreetMap](https://www.openstreetmap.org) `XYZ` tiles. I'll refer to data that use longitude and latitude as _geographic_ data.

## the problem

The problem I'm concerned with is that when you try to compute a Voronoi diagram by _directly_ applying the standard algorithm, e.g., deldir::deldir(), to  _geographic_ data you get the "wrong" result:

```{r wrong, echo = FALSE, fig.align = "center"}
vars <- c("lon", "lat")
asp <- 1.6
cases <- cholera::fatalities.address
rng <- mapRange(latlong = TRUE)
snow.colors <- snowColors(vestry = FALSE)

plot(cases[, vars], xlim = rng$x, ylim = rng$y, pch = NA, asp = asp)
addRoads(vars)
points(pumps[, vars], pch = 17, col = snow.colors, cex = 1)
text(pumps[, vars], pos = 1, labels = pumps$id)
points(cases[, vars], pch = 16, col = "gray", cex = 0.5)
cells <- cholera::voronoiPolygons(cholera::pumps[, vars],
  rw.data = cholera::roads[, vars], latlong = TRUE)
invisible(lapply(cells, polygon))
title(main = "Figure 2 Unadjusted Geographic")
```

While there's nothing obviously "wrong" with the diagram, I'm going to argue that there are two problems with it. First, it doesn't look like diagram fitted to the nominal data in Figure 1. Second, it actually places some people in the wrong neighborhood, which is something that, by definition, shouldn't happen.

## what should the Voronoi diagram should look like?

The Voronoi diagram should actually look similar to the diagram in Figure 1 even though they're fitted to data with different coordinates, nominal and geographic. To understand why, we need to dig a little deeper into how Voronoi diagrams work. Above, I actually described the intuition behind them: find the closest pump for everyone in town and distinct neighborhoods will emerge. This _pumps-and-people_ approach is the "brute force" way to uncover neighborhoods. However, the implementation behind the standard algorithm actually uses a _pumps-only_ approach to the problem. This is one of the really neat things about the algorithm: you only need the location of water pumps to compute the Voronoi diagram (it's also makes algorithm efficient).

Because of this, the geographic Voronoi diagram _should_ look the nominal one. After all, as we can see in Figures 4 & 5, which exclude the Voronoi diagrams, the relative positions of the pumps for the nominal and geographic data are quite similar:

```{r nominal_geographic, echo = TRUE, fig.align = "default", fig.width = 5.5, fig.height = 5.5, fig.show = "hold", out.width = "50%"}
cases <- cholera::fatalities.address
snow.colors <- snowColors(vestry = FALSE)

vars <- c("x", "y")
rng <- mapRange()
asp <- 1
plot(cases[, vars], xlim = rng$x, ylim = rng$y, pch = NA, asp = asp)
addRoads()
points(pumps[, vars], pch = 17, col = snow.colors, cex = 1)
text(pumps[, vars], pos = 1, labels = pumps$id)
points(cases[, vars], pch = 16, col = "gray", cex = 0.5)
title(main = "Figure 4 Nominal")

vars <- c("lon", "lat")
rng <- mapRange(latlong = TRUE)
asp <- 1.6
plot(cases[, vars], xlim = rng$x, ylim = rng$y, pch = NA, asp = asp)
addRoads(vars)
points(pumps[, vars], pch = 17, col = snow.colors, cex = 1)
text(pumps[, vars], pos = 1, labels = pumps$id)
points(cases[, vars], pch = 16, col = "gray", cex = 0.5)
title(main = "Figure 5 Geographic")
```

As a result, the graph should look like this:

```{r adjusted, echo = FALSE, fig.align = "center"}
plot(cases[, vars], xlim = rng$x, ylim = rng$y, pch = NA, asp = asp)
addRoads(vars)
points(pumps[, vars], pch = 17, col = snow.colors, cex = 1)
text(pumps[, vars], pos = 1, labels = 1:13)
points(cases[, vars], pch = 16, cex = 0.5, col = "gray")
cells <- cholera::latlongVoronoiC()
invisible(lapply(cells, function(dat) polygon(dat[, vars])))
title(main = "Figure 6 Adjusted Geographic")
```

As we'll see in the next section, that it looks like Figure 1 is no coincidence.

### "brute force" pumps-and-people method behind the intuition

While it may not be computationally efficient, the value of the "brute force" _pumps-and-people_ approach goes beyond its ability to illustrate the underlying intuition. For me, its value lies with the fact that it provides a separate, independent way to compute and validate a Voronoi diagram

As "proof" of the equivalence of the _pumps-and-people_ (intuition) and the _pumps-only_ (implementation) methods, compare the _pumps-only_ diagram in Figure 1 with the illustration of the "brute force" method in Figure 6:

```{r euclidean_paths, echo = FALSE, fig.align = "center"}
plot(neighborhoodEuclidean(case.set = "expected"), type = "star",
  add.title = FALSE)
title(main = "Figure 7 Brute Force Pumps-and-People Voronoi Diagram")
```

I created this graph by placing 20,000 regularly spaced points across the face of the map. Next, I computed the distance from each point to the 13 water pumps. Then, I drew a line from each point to its closest pump (color-coded by pump). What emerges is evidence of equivalence in the form of a kind of photographic negative or inverse of the pump-only diagram.

With this tool in hand, I'll show that the "Unadjusted Geographic" graph in Figure 2 is "wrong" because it places some people in the wrong neighborhood, which is something that shouldn't happen by definition.

## classification error

To show that the diagram in Figure 2 mis-classifies some people, we need to implement the "brute force" approach for geographic data. To do that we need a way compute to compute distances between pumps and people using geographic coordinates. I do so by using the geosphere::distGeo() function to compute the the geodesic or great circle distance. This is the distance between geographic coordinates along the curved surface of the Earth as determined by the WGS84 ellipsoid model.

The key assumption here is that these geodesic distances represent the "truth on the ground". As such, they serve as the point of reference to compare diagrams in Figures 2 and 3. Figures 7 and 8 re-plot those diagrams and add the independently computed geodesic distances as color-coded line segments that connect a person to their nearest pump. Note that in contrast to the "brute force" plot in Figure 6, I'm only using the 321 unique locations in the data that had at least one fatality:

```{r brute_force}
# "brute force" computation of geodesic distance to nearest pump

cases <- cholera::fatalities.address

nearest.pump <- do.call(rbind, lapply(cases$anchor, function(x) {
  p1 <- cases[cases$anchor == x, vars]
  d <- vapply(pumps$id, function(p) {
    p2 <- pumps[pumps$id == p, vars]
    geosphere::distGeo(p1, p2)
  }, numeric(1L))
  nearest <- which.min(d)
  data.frame(case = x, pump = pumps$id[pumps$id == nearest],
    meters = d[nearest])
}))

```

```{r unadjusted_classification, echo = FALSE, fig.align = "center"}
# variables

vars <- c("lon", "lat")
asp <- 1.6

rng <- mapRange(latlong = TRUE)
snow.colors <- snowColors(vestry = FALSE)

# plot roads and pumps

plot(cases[, vars], xlim = rng$x, ylim = rng$y, pch = NA, asp = asp)
addRoads(vars)
points(pumps[, vars], pch = 17, col = snow.colors, cex = 1)

# compute Voronoi cells by directly applying the algorithm to geographic data

cells <- cholera::voronoiPolygons(cholera::pumps[, vars],
  rw.data = cholera::roads[, vars], latlong = TRUE)

# plot Voronoi cells

invisible(lapply(cells, polygon))

# plot color-coded line segment from fatality to pump

invisible(lapply(nearest.pump$case, function(x) {
  ego <- cases[cases$anchor == x, vars]
  p <- nearest.pump[nearest.pump$case == x, "pump"]
  alter <- pumps[pumps$id == p, vars]
  segments(ego$lon, ego$lat, alter$lon, alter$lat,
    col = snow.colors[paste0("p", p)])
}))

# plot color-coded fatalities

invisible(lapply(nearest.pump$case, function(x) {
  ego <- cases[cases$anchor == x, vars]
  p <- nearest.pump[nearest.pump$case == x, "pump"]
  points(ego, pch = 16, col = snow.colors[paste0("p", p)], cex = 2/3)
}))

title(main = "Figure 8 Undjusted Geographic")
```

```{r adjusted_classification, echo = FALSE, fig.align = "center"}

plot(cases[, vars], xlim = rng$x, ylim = rng$y, pch = NA, asp = asp)
addRoads(vars)
points(pumps[, vars], pch = 17, col = snow.colors, cex = 1)

cells <- cholera::latlongVoronoiC()
invisible(lapply(cells, function(dat) polygon(dat[, vars])))

invisible(lapply(nearest.pump$case, function(x) {
  ego <- cases[cases$anchor == x, vars]
  p <- nearest.pump[nearest.pump$case == x, "pump"]
  alter <- pumps[pumps$id == p, vars]
  segments(ego$lon, ego$lat, alter$lon, alter$lat, lwd = 0.5,
    col = snow.colors[paste0("p", p)])
}))

invisible(lapply(nearest.pump$case, function(x) {
  ego <- cases[cases$anchor == x, vars]
  p <- nearest.pump[nearest.pump$case == x, "pump"]
  points(ego, pch = 16, col = snow.colors[paste0("p", p)], cex = 0.5)
}))

title(main = "Figure 9 Adjusted Geographic")
```

In the "Unadjusted Geographic" graph (Figure 8), you can see discrepancies between the cells and a person's closest pump, which by definition shouidn't occur. In the "Adjusted Geographic" graph (Figure 9) there are no discrepancies.

## a solution

How does this adjustment in Figure 8 work? I use a familiar strategy: I transform the data so that I _can_ use the standard algorithm; I then translate the results back to their original form (longitude and latitude).

First, I transform the geographic coordinates so that they represent relative rather than absolute positions. I do so by computing the geodesic distance from an arbitrary point of reference (a "corner of the map") to the 13 water pumps. Second, I break down those distances into its horizontal and vertical components, _meters-North_ and _meters-East_ of the reference point. These components, the analogs of latitude and longitude, serve as our new set of coordinates. The virtues of these new coordinates is that they are Cartesian coordinates, which the standard algorithm "expects", and that they allow us to apply the standard algorithm and to get the "right" results (Figure 8). Third, I translate the results, which are the coordinates of the cell vertices, back to longitude and latitude. To do that, I use data simulation to uncover the relationship between geodesic distance and longitude, and the relationship between geodesic distance and latitude. While both relationships, for Soho at least, could statistically speaking be summarized with a linear model (OLS), I instead fit both to separate loess functions. I then use those functions to translate the data back to longitude and latitude.

## spatial data approach via 'terra'

Finally, as a sanity check and as additional piece of evidence that I may be on the right track, I replicate the "Unadjusted" and "Adjusted" diagrams using the [spatial](https://rspatial.org/) approach employed in the ['terra'](https://cran.r-project.org/package=terra) package.

```{r, terra_data, fig.align = "default", fig.width = 5.5, fig.height = 5.5, fig.show = "hold", out.width = "50%", fig.cap = "Figure 7 Nominal Analog with 'terra' package"}
dat <- cholera::pumps[, c("lon", "lat")]
sv.data <- terra::vect(dat, crs = "+proj=longlat")
v1 <- terra::voronoi(sv.data)

pt1 <- "+proj=lcc +lat_1=51.510 +lat_2=51.516"
pt2 <- "+lat_0=51.513 +lon_0=-0.1367 +units=m"
proj <- paste(pt1, pt2)
sv.proj <- terra::project(sv.data, proj)
v2 <- terra::voronoi(sv.proj)

out1 <- terra::project(v1, "+proj=longlat")
out2 <- terra::project(v2, "+proj=longlat")

plot(out1, xlim = range(cholera::roads$lon), ylim = range(cholera::roads$lat))
addRoads(c("lon", "lat"))
points(cholera::pumps[, c("lon", "lat")])
text(cholera::pumps[, c("lon", "lat")], labels = 1:13, pos = 1)
title(main = "Figure 10 Undjusted Geographic")

plot(out2, xlim = range(cholera::roads$lon), ylim = range(cholera::roads$lat))
addRoads(c("lon", "lat"))
points(cholera::pumps[, c("lon", "lat")])
text(cholera::pumps[, c("lon", "lat")], labels = 1:13, pos = 1)
title(main = "Figure 11 Adjusted Geographic")
```

## appendix: solution

The following walks through the code for the working solution, which can be found in [latlongVoronoiC.R](https://github.com/lindbrook/cholera/blob/master/R/latlongVoronoiC.R)

#### four corners

Origin is bottom left; graph is in quadrant I

```{r four_corners}
origin <- data.frame(lon = min(cholera::roads$lon),
                     lat = min(cholera::roads$lat))
topleft <- data.frame(lon = min(cholera::roads$lon),
                      lat = max(cholera::roads$lat))
bottomright <- data.frame(lon = max(cholera::roads$lon),
                          lat = min(cholera::roads$lat))
topright <- data.frame(lon = max(cholera::roads$lon),
                       lat = max(cholera::roads$lat))
```

#### break down geodesic distance into horizontal and vertical components

Compute geodesic distance from origin to points and decompose

```{r decomposed_coordinates}
pump.data <- cholera::pumps

pump.meters <- do.call(rbind, lapply(pump.data$id, function(p) {
  pmp <- pump.data[pump.data$id == p, c("lon", "lat")]
  x.proj <- c(pmp$lon, origin$lat)
  y.proj <- c(origin$lon, pmp$lat)
  m.lon <- geosphere::distGeo(y.proj, pmp)
  m.lat <- geosphere::distGeo(x.proj, pmp)
  data.frame(pump = p, x = m.lon, y = m.lat)
}))
```

#### bounding box of Voronoi diagram

```{r, bounding_box}
height <- geosphere::distGeo(origin, topleft)
width <- geosphere::distGeo(origin, bottomright)
bounding.box <- c(0, width, 0, height)
```

#### cell coordinates

Apply deldir::deldir() and extract coordinates of cells, for use with polygon().

```{r, cell_coords}
cells <- voronoiPolygons(pump.meters[, c("x", "y")], rw = bounding.box)
```

#### reshape and reformat into data frame

```{r, format data}
cells.df <- do.call(rbind, cells)
cells.lat <- sort(unique(cells.df$y), decreasing = TRUE) # unique latitudes
tmp <- row.names(cells.df)
ids <- do.call(rbind, strsplit(tmp, "[.]"))
cells.df$cell <- as.numeric(ids[, 2])
cells.df$vertex <- as.numeric(ids[, 3])
row.names(cells.df) <- NULL
```

#### translating back to longitude and latitude

To translate the results (the coordinates of Voronoi cells) from meters-East and meters-North back to longitude and latitude, I use data simulation to uncover the relationship between geodesic distance and units of longitude and latitude.

While the measure of geodesic distance I use, geosphere::distGeo(), use an ellipsoid model of the Earth (WGS 84), the relationship between meters-North and latitude is perfectly linear for Soho, Westminster (UK). Nevertheless, I play it conservatively and use a fitted loess function to translate meters-North to latitude. The details are in the meterLatitude():

```{r, meterLatitude}
meterLatitude <- function(cells.df, origin, topleft, delta = 0.000025) {
  lat <- seq(origin$lat, topleft$lat, delta)

  meters.north <- vapply(lat, function(y) {
    geosphere::distGeo(origin, cbind(origin$lon, y))
  }, numeric(1L))

  loess.lat <- stats::loess(lat ~ meters.north,
    control = stats::loess.control(surface = "direct"))

  y.unique <- sort(unique(cells.df$y))

  est.lat <- vapply(y.unique, function(m) {
    stats::predict(loess.lat, newdata = data.frame(meters.north = m))
  }, numeric(1L))

  data.frame(m = y.unique, lat = est.lat)
}
```

East-West (horizontal) distance is a function of how far North or South (latitude) you are. This means that the relationship between meters-East and longitude will be nonlinear. While analysis indicates that a fitted OLS line might be adequate, I fit separate loess functions for each observed meters-North value. The details are in the meterLatLong():

```{r meterLatLong}
meterLatLong <- function(cells.df, origin, topleft, bottomright,
  delta = 0.000025) {

  est.lat <- meterLatitude(cells.df, origin, topleft)

  # uniformly spaced points along x-axis (longitude)
  lon <- seq(origin$lon, bottomright$lon, delta)

  # a set of horizontal distances (East-West) for each estimated latitude
  meters.east <- lapply(est.lat$lat, function(y) {
    y.axis.origin <- cbind(origin$lon, y)
    vapply(lon, function(x) {
      geosphere::distGeo(y.axis.origin, cbind(x, y))
    }, numeric(1L))
  })

  loess.lon <- lapply(meters.east, function(m) {
    dat <- data.frame(lon = lon, m)
    stats::loess(lon ~ m, data = dat,
      control = stats::loess.control(surface = "direct"))
  })

  y.unique <- sort(unique(cells.df$y))

  # estimate longitudes, append estimated latitudes
  est.lonlat <- do.call(rbind, lapply(seq_along(y.unique), function(i) {
    dat <- cells.df[cells.df$y == y.unique[i], ]
    loess.fit <- loess.lon[[i]]
    dat$lon <- vapply(dat$x, function(x) {
      stats::predict(loess.fit, newdata = data.frame(m = x))
    }, numeric(1L))
    dat$lat <- est.lat[est.lat$m == y.unique[i], "lat"]
    dat
  }))

  est.lonlat[order(est.lonlat$cell, est.lonlat$vertex), ]
}
```

The code below shows the data that translates the coordinates for the cell for pump 1 from meters-East and meters-North to longitude and latitude.

```{r, translation}
est.lonlat <- meterLatLong(cells.df, origin, topleft, bottomright)
longlat <- split(est.lonlat, est.lonlat$cell)
```

```{r sample_data}
longlat[[1]]
```
