---
title: "deldirPolygons(): Tiles, Triangles and Polygons"
author: "lindbrook"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = ">")
library(cholera)
```

`deldirPolygons()` is a wrapper function that extracts the vertices of 'deldir' Delauny triangles and Dirichelet (Voronoi) tiles for use with functions that rely on polygons. The function returns a list of data frames of vertices. This makes tasks like coloring tiles or triangles or counting cases within tiles or triangles easier.

```{r deldirPolygons, eval = FALSE}
deldirPolygons(sites, rw.data = NULL, rw = NULL, type = "tiles")
```

The functions has four arguments. `sites` is the data frame of the sites or focal points used to do the triangulation or tessellation. `rw.data` (rw = 'rectangular window')is the data frame of secondary data (e.g., fatalities, customers, etc.). It is useful when the range of secondary data exceeds that of the site data. `rw` is a vector of the corners of the rectangular window: xmin, xmax, ymin, ymax. `type` is "tiles" or "triangles".

## Coloring Tiles

```{r coloring, fig.align = "left", fig.width = 5, fig.height = 5, echo = TRUE, eval = TRUE}
# compute vertices of Voronoi tiles
vertices <- deldirPolygons(sites = cholera::pumps, rw.data = cholera::roads)

# define colors, plot map, and color code fatalities
snow.colors <- grDevices::adjustcolor(cholera::snowColors(), alpha.f = 1/3)
cholera::snowMap(add.cases = FALSE)
cholera::addNeighborhoodCases(metric = "euclidean")

# plot color coded polygons
invisible(lapply(seq_along(vertices), function(i) {
  polygon(vertices[[i]], col = snow.colors[[i]])
}))
```

## Counting Observations in Tiles

To count the number of cases within each neighborhood, we can use sp::point.in.polygon().

```{r counting, echo = TRUE, eval = TRUE}
# compute vertices of Voronoi tiles
vertices <- deldirPolygons(sites = cholera::pumps, rw.data = cholera::roads)

# locations of the 578 fatalities in Soho
cases <- cholera::fatalities.unstacked

# count fatalities within each polygon (neigborhood)
census <- lapply(vertices, function(tile) {
  sp::point.in.polygon(cases$x, cases$y, tile$x, tile$y)
})

# ID the 13 water pumps
names(census) <- paste0("p", cholera::pumps$id)

# count of fatalities by neighborhood
vapply(census, sum, integer(1L))
```

## Counting Observations in Triangles

To count the number of cases within each triangle:

```{r counting_triangles, echo = TRUE, eval = TRUE}
# compute vertices of Delauny triangles
vertices <- deldirPolygons(sites = cholera::pumps,
  rw.data = cholera::roads, type = "triangles")

# locations of the 578 fatalities in Soho
cases <- cholera::fatalities.unstacked

# count fatalities within each triangle
census <- lapply(vertices, function(tile) {
  sp::point.in.polygon(cases$x, cases$y, tile$x, tile$y)
})

# ID triangles
names(census) <- paste0("t", seq_along(vertices))

# count of fatalities by triangle
vapply(census, sum, integer(1L))
```
