---
title: "Lab Notes: Voronoi tiles to polygons"
author: "lindbrook"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = ">")
library(cholera)
```

## Voronoi tiles to polygons

Using just the locations of sites or landmarks, Voronoi tessellation partitions a space into cells or tiles that represent those sites' neighborhoods (i.e., their catchment or service areas).

To make tasks like coloring tiles or counting cases within tiles easier, I wrote voronoiPolygons() function. It converts 'deldir' Voronoi tessellation tiles into polygons (a list of data frames of vertices).

## Coloring Polygons

```{r coloring, fig.align = "left", fig.width = 5, fig.height = 5, echo = TRUE, eval = TRUE}
# compute vertices of Voronoi tiles
polygon.vertices <- voronoiPolygons(sites = cholera::pumps,
  rw.data = cholera::roads)

# define colors, plot map, and color code fatalities
snow.colors <- grDevices::adjustcolor(cholera::snowColors(), alpha.f = 1/3)
cholera::snowMap(add.cases = FALSE)
cholera::addNeighborhoodCases(metric = "euclidean")

# plot color coded polygons
invisible(lapply(seq_along(polygon.vertices), function(i) {
  polygon(polygon.vertices[[i]], col = snow.colors[[i]])
}))
```

## Counting Observations in Polygons

To count the number of cases within each neighborhood, we can use sp::point.in.polygon().

```{r counting, echo = TRUE, eval = TRUE}
# compute vertices of Voronoi tiles
polygon.vertices <- voronoiPolygons(sites = cholera::pumps,
  rw.data = cholera::roads)

# locations of the 578 fatalities in Soho
cases <- cholera::fatalities.unstacked

# count fatalities within each polygon (neigborhood)
census <- lapply(polygon.vertices, function(tile) {
  sp::point.in.polygon(cases$x, cases$y, tile$x, tile$y)
})

# ID the 13 water pumps
names(census) <- paste0("p", cholera::pumps$id)

# count of fatalities by neighborhood
vapply(census, sum, integer(1L))
```
